JuliaCon Proceedings in Quarto

### 2023-11-16 template.qmd
this seems to work

---
title: Juliacon Proceedings in Quarto
format:
  juliacon-proceedings-pdf:
    keep-tex: true  
  juliacon-proceedings-html: default
author:
  - name: Patrick Altmeyer
    affiliations:
      - Delft University of Technology
    orcid: 0000-0003-4726-8613
    email: p.altmeyer@tudelft.nl
    url: https://www.paltmeyer.com/
abstract: |
  This document is a template demonstrating the `juliacon-proceedings` format. It is a standard Quarto document composed of Markdown, LaTeX and code chunks. It is rendered to PDF and HTML. The PDF is rendered using the `juliacon-proceedings-pdf` format, which is based on the `juliacon-proceedings` LaTeX class. The HTML is rendered using the `juliacon-proceedings-html` format, which is based on the `juliacon-proceedings` HTML template.
keywords: [Template, Demo, Quarto, JuliaCon]
bibliography: bibliography.bib  
jupyter: julia-1.9
---

@@

## About this template

This template is based on the existing [JuliaCon Proceedings LaTeX template](https://github.com/JuliaCon/JuliaConSubmission.jl).

## How to use this template

> The remainder of this document repeats relevant elements from the [JuliaCon Proceedings LaTeX template].

# The JuliaCon Article Class {#sec-documentclass}

The `juliacon` class file preserves the standard LATEX{} interface such that any document that can be produced using the standard LATEX{} article class can also be produced with the class file.

It is likely that the make up will change after file submission. For this reason, we ask you to ignore details such as slightly long lines, page stretching, or figures falling out of synchronization, as these details can be dealt with at a later stage.

Use should be made of symbolic references (\verb|\ref| ) in order to protect against late changes of order, etc.

# Executable Code

Ideally, we would be able to tab into Quarto's existing support for executable, cross-referenceable code chunks.

@@

For example, the following code block will be labeled as `lst-1` and calling `@lst-1` will render as @lst-1. This ensures that things stay up-to-date and potential errors in the code are identified early: as you render your document, the code will be executed and the output will be inserted into the document.

```{julia}
#| lst-label: lst-1
#| lst-cap: A listing caption
#| echo: true

@@

x = -3.0:0.01:3.0
y = rand(length(x));
```

Unfortunately, I don't know how to use the special environment that is already defined for Julia in this context.

# Code Listings

The special environment that is already defined for Julia code can still be used as before.

@@

>>>>

---
title: JuliaCon Proceedings in Quarto
format:
  juliacon-proceedings-pdf:
    keep-tex: true  
  juliacon-proceedings-html: default
author:
  - name: Patrick Altmeyer
    affiliations:
      - Delft University of Technology
    orcid: 0000-0003-4726-8613
    email: p.altmeyer@tudelft.nl
    url: https://www.paltmeyer.com/
abstract: |
  This document is a template demonstrating the `juliacon-proceedings` format. It is a standard Quarto document composed of Markdown, LaTeX and code chunks. It is rendered to PDF and HTML. The PDF is rendered using the `juliacon-proceedings-pdf` format, which is based on the `juliacon-proceedings` LaTeX class. The HTML is rendered using the `juliacon-proceedings-html` format, which is based on the `juliacon-proceedings` HTML template.
keywords: [Template, Demo, Quarto, JuliaCon]
bibliography: bibliography.bib  
jupyter: julia-1.9
---

@@

# About this template

This template is based on the existing [JuliaCon Proceedings LaTeX template](https://github.com/JuliaCon/JuliaConSubmission.jl). The rendered versions can be found here: [pdf](https://www.paltmeyer.com/quarto-juliacon-proceedings/template.pdf), [html](https://www.paltmeyer.com/quarto-juliacon-proceedings/template.html).

## Basic Usage

To use the Quarto [extension](https://github.com/pat-alt/quarto-juliacon-proceedings) that provides this template, you can create a new project as follows:

```
quarto use template pat-alt/quarto-juliacon-proceedings
```

Alternatively, you can add the extension to an existing project:

```
quarto add pat-alt/quarto-juliacon-proceedings
```

Then, add the format to your document options:

```yaml
format:
  juliacon-proceedings-pdf: default
```

# Code

There are various options to present code using this template. We could leverage Quarto's existing support for executable, cross-referenceable code. Alternatively, we could use the existing `lstlisting` environment for Julia code.

## Executable Listings  (version $\geq$ 1.4.0)

Relying on Quarto's support for executable listings is the most flexible option as it allows us to use all of Quarto's features. It comes with numerous advantages:

1. We can use the same code chunks for both PDF and HTML output (and more) with consistent formatting.
2. Using executable code ensures that things stay up-to-date and potential errors in the code are identified early: as you render your document, the code will be executed and the output will be inserted into the document.
3. The output of executable code can be cross-referenced. For example, we can reference a figure that is generated by a code chunk and the reference will be updated automatically when the figure number changes.
4. Code chunks can also be hidden. In HTML, there is additional support for code folding.
5. The code itself can be cross-referenced.

From the [Quarto Docs](https://quarto.org/docs/prerelease/1.4/crossref.html#cross-referenceable-listings-of-executable-code-blocks):

@@

For example, the following code block will be labeled as `lst-1` and calling `@lst-1` will render as @lst-1. The figure it generates will be labeled as `fig-1` and calling `@fig-1` will render as @fig-1.

```{julia}
#| lst-label: lst-1
#| lst-cap: A listing caption.
#| label: fig-1
#| fig-cap: A figure caption.
#| echo: true

@@

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
```

::: {.callout-note}

# Current Limitations

Support for executable listings is only available in the development version of Quarto (version $\geq$ 1.4.0). It is currently not clear how to support custom formatting, for example through the existing `jlcode.sty` environment.

:::

## Static Listings

The special environment that is already defined for Julia code is still compatible with this template.

@@

# Tables

As with code, there are various options to present tables. Using standard Markdown syntax is the most flexible option because it is compatible with various output formats. Alternatively, standard LaTeX syntax can be used.

## Markdown Tables

Markdown tables are rendered to LaTeX using the `booktabs` package. This is the recommended way to create tables because it is compatible with various output formats. Tables can be cross-referenced and the table caption will be rendered as a LaTeX table caption. The table generated through the following Markdown syntax is labeled as `tbl-1` and calling `@tbl-1` will render as @tbl-1.

```{.markdown}

| Default | Left | Right | Center |
|---------|:-----|------:|:------:|
| 12      | 12   |    12 |   12   |
| 123     | 123  |   123 |  123   |
| 1       | 1    |     1 |   1    |

: Demonstration of pipe table syntax {#tbl-1}

```

| Default | Left | Right | Center |
|---------|:-----|------:|:------:|
| 12      | 12   |    12 |   12   |
| 123     | 123  |   123 |  123   |
| 1       | 1    |     1 |   1    |

: Demonstration of pipe table syntax {#tbl-1}
